### 2017-07-07 Plos-one-manuscript/latex/manuscript.tex
save

\section*{Methods}
Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer (fig.~\ref{fig1}), in this case a Makerbot Replicator.
A 1 mL or 3 mL syringe twists to lock in the body part and is held into place by the syringe flanges.
The 30-300 $\mu$L configuration uses a 1 mL pipette and the 100-1000 $\mu$L configuration uses a 3 mL pipette.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb button (fig.~\ref{fig2}). 
The pipette is spring loaded towards a set point which is adjustable by a set-screw.
When the thumb button is pressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with the unlatching button drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to `blow-out' the transferred fluid completely from the pipette tip.   
Additional materials required for assembly include two springs, a nut, a bolt and two washers.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead, a combination of a barbed luer adapter and elastic tubing is used to attach the pipette tips.
Our printed pipette mimics commercial pipettes design, function, and user operation, making it intuitive to use.

@@

\subsection*{Fabrication \& Assembly}
Two parts are printed (body.stl and plunger.stl) at normal resolution with rafts on a Makerbot Replicator.
A small amount of paraffin wax is applied to the screw to prevent slop from causing the set point to drift after each actuation.
The nut is sunk into the hex inset in the printed body part.
The bolt is threaded in from the top of the body into the nut.
Springs and washers are threaded onto the plunger of a 1 mL syringe for the 30-300 $\mu$L configuration.
Springs are placed inside a 3 mL syringe for the 100-1000 $\mu$L configuration.
The plunger part is inserted in the body and the syringe assembly is pushed in and locked from the syringe flanges to the body part to complete assembly(fig.~\ref{fig3}).
Parts and cost are listed in Table~\ref{table1} and \ref{table2}.

@@

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 30-300 $\mu$L pipette.}
\label{table1}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	1 mL Syringe & \$0.15     & BD Biosciences & 309628      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	Washers (2)  & \$0.16     & McMaster-Carr  & 90107A012   \\ \hline
	total        & \$6.21     &                &             \\ \hline
\end{tabular}
\end{table}

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 100-1000 $\mu$L pipette.}
\label{table2}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	3 mL Syringe & \$0.73     & BD Biosciences & 309657      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	total        & \$6.63     &                &             \\ \hline
\end{tabular}
\end{table}

@@

Our solution was to replace the built-in scale with a scale that accounted for the expansion.
We compared expected volumes with actual measurements and noticed a linear relationship.
We scaled up the graduations by a factor of 1.027x for both the 100-1000 $\mu$L, and 30-300 $\mu$L configurations.
These scales were exported to portable document format (PDF) so that they are most easily printed to scale.
These adjusted scales are printed on paper or a transparency sheet and taped onto the syringe with the zero at plunger in the latched position (\nameref{S1_Fig} and \nameref{S1_Fig}).

>>>>

\section*{Methods}
Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer (fig.~\ref{fig1}), in this case a Makerbot Replicator.
A 1 mL or 3 mL syringe twists to lock in the body part and is held into place by the syringe flanges.
The 30-300 $\mu$L configuration uses a 1 mL pipette and the 100-1000 $\mu$L configuration uses a 3 mL pipette.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb button (fig.~\ref{fig2}). 
The pipette is spring loaded towards a set point which is adjustable by a set-screw.
When the thumb button is pressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with the unlatching button drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to `blow-out' the transferred fluid completely from the pipette tip.   
Additional materials required for assembly include two springs, a nut, and a bolt.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead, a combination of a barbed luer adapter and elastic tubing is used to attach the pipette tips.
Our printed pipette mimics commercial pipettes design, function, and user operation, making it intuitive to use.

@@

\subsection*{Fabrication \& Assembly}
Two parts are printed (body.stl and plunger.stl) at normal resolution with rafts on a Makerbot Replicator.
A small amount of paraffin wax is applied to the screw to prevent slop from causing the set point to drift after each actuation.
The nut is sunk into the hex inset in the printed body part.
The bolt is threaded in from the top of the body into the nut.
Two springs are threaded onto the plunger of a 1 mL syringe for the 30-300 $\mu$L configuration.
Springs are placed inside a 3 mL syringe for the 100-1000 $\mu$L configuration.
The plunger part is inserted in the body and the syringe assembly is pushed in and locked from the syringe flanges to the body part to complete assembly(fig.~\ref{fig3}).
Parts and cost are listed in Table~\ref{table1} and \ref{table2}.

@@

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 30-300 $\mu$L pipette.}
\label{table1}
	\begin{tabular}{|l|l|r|l|}
		\hline
	30-300 mL Configuration:              & Source            & Unit Price & Part Number \\ \hline
	Filament                              & Makerbot          & \$1.63     & NA          \\ \hline
	1 mL Syringe                          & BD Biosciences    & \$0.15     & 309628      \\ \hline
	M3 Bolt, 35 mm                        & McMaster-Carr     & \$0.12     & 91287A026   \\ \hline
	M3 Nut                                & McMaster-Carr     & \$0.01     & 90591A121   \\ \hline
	Music Wire Compression Springs* (2)   & Jones Spring Co   & \$1.23     & C10-022-048 \\ \hline
	Female Luer to 1/8" Hose Barb Adapter & Cole-Parmer       & \$0.40     & EW-30800-08 \\ \hline
	Tygon Tubing, 1/16"ID x 3/16"OD       & Cole-Parmer       & \$0.03     & EW-06407-72 \\ \hline
	300 uL Tips (10)                      & Fisher Scientific & \$0.34     & 02-707-447  \\ \hline
	& Total             & \$3.91     &     \\ \hline        
	\end{tabular}
\end{table}

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 100-1000 $\mu$L pipette.}
\label{table2}
\begin{tabular}{|l|l|r|l|}
	\hline
	100-1000 mL configuration:             & Source            & Unit Price & Part Number \\ \hline
	Filament                               & Makerbot          & \$1.63     & NA          \\ \hline
	3 mL Syringe                           & BD Biosciences    & \$0.73     & 309657      \\ \hline
	M3 Bolt, 35 mm                         & McMaster-Carr     & \$0.12     & 91287A026   \\ \hline
	M3 Nut                                 & McMaster-Carr     & \$0.01     & 90591A121   \\ \hline
	Music Wire Compression Springs* (2)    & Jones Spring Co   & \$1.23     & C10-022-048 \\ \hline
	Female luer to 5/32" Hose Barb Adapter & Cole-Parmer       & \$0.66     & EW-45508-06 \\ \hline
	Tygon Tubing, 1/8"ID x 1/4"OD          & Cole-Parmer       & \$0.05     & EW-06407-76 \\ \hline
	1000 uL Tips (10)                      & Fisher Scientific & \$0.41     & 02-707-400  \\ \hline
	& Total             & \$4.84     &            \\ \hline
\end{tabular}
\begin{flushleft} *Spring Specifications: Overall length 38 mm, OD 7.62 mm, wire diameter 0.56 mm, Max load 0.86 kg, with closed and flat ends. Also available from McMaster-Carr P/N: 9657K347.
\end{flushleft}
\end{table}

@@

Our solution was to replace the built-in scale with a scale that accounted for the expansion.
We compared expected volumes with actual measurements and noticed a linear relationship.
We scaled up the graduations by a factor of 1.027x for both the 100-1000 $\mu$L, and 30-300 $\mu$L configurations.
These scales were exported to portable document format (PDF) so that they are most easily printed to scale.
These adjusted scales are printed on paper or a transparency sheet and taped onto the syringe with the zero at the plunger in the latched position (\nameref{S1_Fig} and \nameref{S1_Fig}).
